\protect Connecting every bit of knowledge: \\ The structure of Wikipedia's First Link Network

### 2016-04-22 writeup/wikipedia-network.tex
bibtex typos

\title{\protect
Connecting every bit of knowledge: \\
The structure of Wikipedia's Flrst Link Network
}

@@

\begin{abstract}
  \protect
Apples, porcupines, and the most obscure Bob Dylan song---is every topic a few clicks from Philosophy? 
Within Wikipedia, the surprising answer is yes: nearly all 
paths lead to Philosophy.
Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
More than information about a topic, Wikipedia is a web of naturally emerging relationships.  
By following the first link in each article, we algorithmically construct a directed network of 
all 4.7 million articles: Wikipedia's First Link Network.
Here, we study the English edition of Wikipedia's First Link Network for insight into how the many 
articles on inventions, places, people, objects, and events are related and organized.  
We traverse every path, measuring the accumulation of first links, path lengths, basins, cycles, 
and the influence each article exerts in shaping the network.
We find scale-free distributions describe path length, 
accumulation, and influence. Far from dispersed, first links disproportionately accumulate 
at a few articles---flowing from specific to general and culminating around fundamental notions such as
Community, State, and Science. 
Philosophy funnels more paths than any other article by two orders of magnitude. 
We also observe a gravitation towards topical articles such as 
Health Care and Fossil Fuel. 
These findings enrich our view of the connections and structure of
Wikipedia's ever growing store of knowledge.

@@

%TODO:
% Abstract 
    % In traversing every path, we measure ?
    % Our findings... braided network (what does this mean?)
% Formatting
    % incorporate into template
% Intro
    % add month to monthly view stats
% Method
    % what to cite for drainage basin in geomorphology

% continue on Page 3

\section{Introduction}

Wikipedia is a towering achievement of the Internet age. 
At no point in history has a larger or more meticulously indexed collection of human knowledge 
existed.
Wikipedia contains 37 million articles in 283 languages, 
with coverage spanning everything from little known ancient battles to the latest pharmaceutical drugs 
\cite{drugs, stats}.
Demonstrating its relevance to modern inquiry,
Wikipedia is the sixth most visited
site in the world, surpassing 18 billion page views and 11 million edits in a single month.
\cite{wiki_edits, wiki_views}.

Wikipedia has naturally become the object of many studies. 
Researchers have examined the cultural dynamics among editors
\cite{editors},
the accuracy of the content relative to traditional encyclopedias
\cite{accuracy1, accuracy2},
the topics covered 
\cite{coverage},
and bias against portions of the population
\cite{bias_women}.
Wikipedia's content has also proven to be a powerful tool. 
Researchers have used Wikipedia to identify missing dictionary entries
\cite{missing_entries},
cluster short text
\cite{clustering},
compute semantic relatedness
\cite{semantic_relatedness},
and disambiguate meaning \cite{disambiguating}.

While these many studies have dissected and fruitfully applied Wikipedia's content,
here we examine the connections among the many articles.
A hyperlink from one Wikipedia article to another
naturally indicates a relationship between the two articles
\cite{relevance}.
The notion that hyperlinks convey information about the content of a
page has proved enormously successful in multiple domains from search engine algorithms 
such as PageRank 
\cite{pagerank} 
to topic classification
\cite{classifier}.
Here, we treat a hyperlink as a mechanism connecting two topics.

@@

We also analyze article popularity for the highest ranking 1000 articles by 
traversal funnels. In October, the average page views per article is 
$22$ thousand with a standard deviation of $95$ thousand views. The distribution
is skewed with $75\%$ of articles reaching fewer than $20$ thousand page views. 
``Halloween'' is the most popular article with $2.8$ million views in October,
likely a result of Halloween falling in the month of October. 
Other popular articles include ``Scientology'' ($463$ thousand views), ``Clint Eastwood'' 
($341$ thousand views),
and the ``Cold War'' ($298$ thousand views) although each has significantly fewer views compared to the views for ``Halloween''.
Other standout popular October articles include ``24-hour'' and ``12-hour'' clocks likely due
to Daylight savings in October and ``Marriage'' possibly due to the drop in the number of
weddings in the months following October
\cite{weddings}.
``Philosophy'' which ranks seventh among the most popular of the highest ranking articles by traversal funnels, appears nowhere in the top 20
by traversal visits. ``Philosophy'' is a relatively popular article 
among articles influencing the shape of the network, but less popular among 
highly-ranked articles by accumulation.

@@

The findings here should only be considered within the limitations of their context.
We examined only the English edition of Wikipedia at a particular moment in time, 
considering only the first link in in each article as a means to relate articles.
Finally, Wikipedia, while the largest digital 
collection of human knowledge, is rife with the biases of the many contributing editors
\cite{bias}.
Nevertheless, the findings do reveal
generalizable relationships and point to foundational notions.

@@

\begin{thebibliography}{1}

      \bibitem{editors} Iba, Takashi, et al. ``Analyzing the creative editing behavior of Wikipedia editors: Through dynamic social network analysis.'' Procedia-Social and Behavioral Sciences 2.4 (2010): 6441-6456.

      \bibitem{accuracy1} Holman Rector, Lucy. ``Comparison of Wikipedia and other encyclopedias for accuracy, breadth, and depth in historical articles.'' Reference services review 36.1 (2008): 7-22.

      \bibitem{accuracy2} Giles, Jim. ``Internet encyclopaedias go head to head.'' Nature 438.7070 (2005): 900-901.

       \bibitem{coverage} Halavais, Alexander, and Derek Lackaff. ``An analysis of topical coverage of Wikipedia.'' Journal of Computer Mediated Communication 13.2 (2008): 429-440.

        \bibitem{bias_women} Hill, Benjamin Mako, and Aaron Shaw. ``The Wikipedia gender gap revisited: characterizing survey response bias with propensity score estimation.'' PloS one 8.6 (2013): e65782.

      \bibitem{missing_entries} Williams, Jake Ryland, Clark, Eric M., Bagrow, James P., Danforth, Christopher M. \& Dodds, Peter Sheridan (2015). Identifying missing dictionary entries with frequency-conserving context models. Phys. Rev. E, 92, 042808.

      \bibitem{clustering} Banerjee, Somnath, Krishnan Ramanathan, and Ajay Gupta. ``Clustering short texts using wikipedia.'' Proceedings of the 30th annual international ACM SIGIR conference on Research and development in information retrieval. ACM, 2007.

      \bibitem{semantic_relatedness} Gabrilovich, Evgeniy, and Shaul Markovitch. ``Computing Semantic Relatedness Using Wikipedia-based Explicit Semantic Analysis.'' IJCAI. Vol. 7. 2007.

      \bibitem{disambiguating} Cucerzan, Silviu. ``Large-Scale Named Entity Disambiguation Based on Wikipedia Data.'' EMNLP-CoNLL. Vol. 7. 2007.

      \bibitem{drugs} Clauson, Kevin A., et al. ``Scope, completeness, and accuracy of drug information in Wikipedia.'' Annals of Pharmacotherapy 42.12 (2008): 1814-1821.

      \bibitem{stats} Wikipedia contributors. ``Wikipedia:Statistics.'' Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 9 Nov. 2015. Web. 13 Nov. 2015.

      \bibitem{pagerank} Page, Lawrence, et al. ``The PageRank citation ranking: bringing order to the Web.'' (1999).

      \bibitem{classifier} Chakrabarti, Soumen, Mukul Joshi, and Vivek Tawde. ``Enhanced topic distillation using text, markup tags, and hyperlinks.'' Proceedings of the 24th annual international ACM SIGIR conference on Research and development in information retrieval. ACM, 2001.

      \bibitem{relevance} Kamps, Jaap, and Marijn Koolen. ``Is Wikipedia link structure different?.'' Proceedings of the second ACM international conference on Web search and data mining. ACM, 2009.

      \bibitem{wiki_train} Wikipedia contributors. ``Train.'' Wikipedia, The Free Encyclopedia. Wikipedia, The Free Encyclopedia, 
          \url{https://en.wikipedia.org/wiki/Train}

    \bibitem{locke} Bolton, Martha Brandt. ``The Taxonomy of Ideas in Locke\textquotesingle s {\it Essay},'' The Cambridge Companion to Locke's ``Essay Concerning Human Understanding.'' Ed.Lex Newman.. 1st ed. Cambridge: Cambridge University Press, 2007. 67-100. Cambridge Companions Online. Web. 13 November 2015. http://dx.doi.org/10.1017/CCOL0521834333.004.

    \bibitem{aristotle} Studtmann, Paul, ``Aristotle's Categories,'' The Stanford Encyclopedia of Philosophy (Summer 2014 Edition), Edward N. Zalta (ed.), \url{http://plato.stanford.edu/archives/sum2014/entries/aristotle-categories/}

    \bibitem{descartes} Smith, Kurt, ``Descartes' Theory of Ideas,'' The Stanford Encyclopedia of Philosophy (Spring 2014 Edition), Edward N. Zalta (ed.), \url{http://plato.stanford.edu/archives/spr2014/entries/descartes-ideas/}

    \bibitem{hist_thesaurus} What is the Historical Thesaurus of the OED - Oxford English Dictionary (Oxford English Dictionary)
    http://public.oed.com/historical-thesaurus-of-the-oed/what-is-the-historical-thesaurus-of-the-oed/

    \bibitem{vacc}  Vermont Advanced Computing Core, \url{http://www.uvm.edu/vacc}.

    \bibitem{wiki_views} Page Views for Wikipedia, Non-mobile site, Normalized (Page Views for Wikipedia, Non-mobile site, Normalized)
    \url{https://stats.wikimedia.org/EN/TablesPageViewsMonthly.htm}

    \bibitem{weddings} Wedding statistics in the United States (Wedding statistics in the United States)
    \url{http://www.soundvision.com/article/wedding-statistics-in-the-united-states}

    \bibitem{six_degrees} Elmacioglu, Ergin, and Dongwon Lee. ``On six degrees of separation in DBLP-DB and more.'' ACM SIGMOD Record 34.2 (2005): 33-40.

    \bibitem{wordnet} Princeton University ``About WordNet.'' WordNet. Princeton University. 2010. 
    \url{http://wordnet.princeton.edu}

    \bibitem{newman} Newman, Mark EJ. ``The structure and function of complex networks.'' SIAM review 45.2 (2003): 167-256.

    \bibitem{disease} Eubank, Stephen, et al. ``Modelling disease outbreaks in realistic urban social networks.'' Nature 429.6988 (2004): 180-184.

    \bibitem{social_nets} Newman, Mark EJ, Duncan J. Watts, and Steven H. Strogatz. ``Random graph models of social networks.'' Proceedings of the National Academy of Sciences 99.suppl 1 (2002): 2566-2572.

    \bibitem{geo_basins} Horton, Robert E. ``Erosional development of streams and their drainage basins; hydrophysical approach to quantitative morphology.'' Geological society of America bulletin 56.3 (1945): 275-370.
    APA

    \bibitem{dodds} Dodds, Peter Sheridan, and Daniel H. Rothman. ``Unified view of scaling laws for river networks.'' Physical Review E 59.5 (1999): 4865.

    \bibitem{food_webs} Garlaschelli, Diego, Guido Caldarelli, and Luciano Pietronero. ``Universal scaling relations in food webs.'' Nature 423.6936 (2003): 165-168.

    \bibitem{mat_blog} brain of mat kelcey (brain of mat kelcey). 
    ``Do all first links lead to Philosophy?''
    \url{http://matpalm.com/blog/2011/08/13/wikipedia-philosophy/}

    \bibitem{Ilmari_first_links} User:Ilmari Karonen/First link (Wikipedia)
    \url{https://en.wikipedia.org/wiki/User:Ilmari_Karonen/First_link}

    \bibitem{xkcd} xkcd: Extended Mind (xkcd: Extended Mind)
    \url{http://xkcd.com/903/}

    \bibitem{reddit} Almost every Wikipedia page links to Philosophy if you keep clicking the first link in every article. 
    \url{https://www.reddit.com/r/InternetIsBeautiful/comments/35asgt/almost_every_wikipedia_page_links_to_philosophy/}

    \bibitem{bias} Wagner, Claudia, et al. ``It's a Man's Wikipedia? Assessing Gender Inequality in an Online Encyclopedia.'' arXiv preprint arXiv:1501.06307 (2015).

    \bibitem{wiki_edits} Wikipedia:Statistics 
    \url{https://en.wikipedia.org/wiki/Wikipedia:Statistics}

    \bibitem{media_wiki_templates} MediaWiki: ``Help: Templates''
    \url{https://www.mediawiki.org/wiki/Help:Templates}

    \bibitem{links_that_speak} Ronen, Shahar, Bruno Gonçalves, Kevin Z. Hu, Alessandro Vespignani, Steven Pinker, and César A. Hidalgo. ``'Links That Speak: The Global Language Network and Its Association with Global Fame.'' Proceedings of the National Academy of Sciences Proc Natl Acad Sci USA 111.52 (2014).

\end{thebibliography}

>>>>

\title{\protect
Connecting every bit of knowledge: \\
The structure of Wikipedia's First Link Network
}

@@

\begin{abstract}
  \protect
Apples, porcupines, and the most obscure Bob Dylan song---is every topic a few clicks from Philosophy? 
Within Wikipedia, the surprising answer is yes: nearly all 
paths lead to Philosophy.
Wikipedia is the largest, most meticulously indexed collection of human knowledge ever amassed. 
More than information about a topic, Wikipedia is a web of naturally emerging relationships.  
By following the first link in each article, we algorithmically construct a directed network of 
all 4.7 million articles: Wikipedia's First Link Network.
Here, we study the English edition of Wikipedia's First Link Network for insight into how the many 
articles on inventions, places, people, objects, and events are related and organized.  
By traversing every path, we measure the accumulation of first links, path lengths, basins, cycles, 
and the influence each article exerts in shaping the network.
We find scale-free distributions describe path length, 
accumulation, and influence. Far from dispersed, first links disproportionately accumulate 
at a few articles---flowing from specific to general and culminating around fundamental notions such as
Community, State, and Science. 
Philosophy directs more paths than any other article by two orders of magnitude. 
We also observe a gravitation towards topical articles such as 
Health Care and Fossil Fuel. 
These findings enrich our view of the connections and structure of
Wikipedia's ever growing store of knowledge.

@@

%TODO:
% Formatting
    % incorporate into template
% Abstract 
    % Our findings... braided network (not sure I understand what this means? braided networks don't come up in the paper)
        % is the idea to say "we developed an algorithm that's generalizable to a broader class of networks?"
% Method
    % what to cite for drainage basin in geomorphology
    % clarify basin discussion of method
    % change name from traversal funnels to non-cycle traversal visits?
% Results
    % run MLE; what do i compare against?
\section{Introduction}

Wikipedia is a towering achievement of the Internet age. 
At no point in history has a larger or more meticulously indexed collection of human knowledge 
existed.
Wikipedia contains 37 million articles in 283 languages, 
with coverage spanning everything from little known ancient battles to the latest pharmaceutical drugs 
\cite{clauson2008scope, stats}.
Demonstrating its relevance to modern inquiry,
Wikipedia is the sixth most visited
site in the world, surpassing 18 billion page views and 10 million edits in January, 2013 alone.
\cite{wiki_edits, wiki_views}.

Wikipedia has naturally become the object of many studies. 
Researchers have examined the cultural dynamics among editors
\cite{iba2010analyzing, samoilenko2016linguistic},
the accuracy of the content relative to traditional encyclopedias
\cite{holman2008comparison, giles2005internet},
the topics covered 
\cite{halavais2008analysis},
and bias against portions of the population
\cite{hill2013wikipedia}.
Wikipedia's content has also proven to be a powerful tool. 
Researchers have used Wikipedia to identify missing dictionary entries
\cite{williams2015identifying},
cluster short text
\cite{banerjee2007clustering},
compute semantic relatedness
\cite{gabrilovich2007computing},
and disambiguate meaning \cite{cucerzan2007large}.

While these many studies have dissected and fruitfully applied Wikipedia's content,
here we examine the connections among the many articles.
A hyperlink from one Wikipedia article to another
naturally indicates a relationship between the two articles
\cite{kamps2009wikipedia}.
The notion that hyperlinks convey information about the content of a
page has proved enormously successful in multiple domains from search engine algorithms 
such as PageRank 
\cite{page1999pagerank} 
to topic classification
\cite{chakrabarti2001integrating}.
Here, we treat a hyperlink as a mechanism connecting two topics.

@@

We also analyze article popularity for the highest ranking 1000 articles by 
traversal funnels. In October, the average page views per article is 
$22$ thousand with a standard deviation of $95$ thousand views. The distribution
is skewed with $75\%$ of articles reaching fewer than $20$ thousand page views. 
``Halloween'' is the most popular article with $2.8$ million views in October,
likely a result of Halloween falling in the month of October. 
The second most popular article, ``Scientology'' ($463$ thousand views) appeared in popular news as actress Leah Remini announced she's leaving the Church of Scientology 
\cite{scientology}.
Other popular articles include ``Clint Eastwood'' ($341$ thousand views),
and the ``Cold War'' ($298$ thousand views) although each has significantly fewer views compared to the views for ``Halloween''.
Other standout popular October articles include ``24-hour'' and ``12-hour'' clocks likely due
to Daylight savings in October and ``Marriage'' possibly due to the drop in the number of
weddings in the months following October
\cite{weddings}.
``Philosophy'' which ranks seventh among the most popular of the highest ranking articles by traversal funnels, appears nowhere in the top 20
by traversal visits with only $240$ thousand page views in October, 2015. ``Philosophy'' is a relatively popular article 
among articles influencing the shape of the network, but less popular among 
highly-ranked articles by accumulation.

@@

The findings here should only be considered within the limitations of their context.
We examined only the English edition of Wikipedia at a particular moment in time, 
considering only the first link in in each article as a means to relate articles.
Finally, Wikipedia, while the largest digital 
collection of human knowledge, is rife with the biases of the many contributing editors
\cite{wagner2015s}.
Nevertheless, the findings do reveal
generalizable relationships and point to foundational notions.

@@

\bibliography{wikipedia-network}{}
\bibliographystyle{plain}
